---
title: "Correlación tamaño - zona censal"
output: 
---

On Page 10, Lines 27-34 you implicitly refer to the Modifiable Areal Unit Problem (MAUP), which is far from a minor issue with respect to the estimation of segregation indices. Perhaps you might consider performing a regression analysis to evaluate the effect of census tract size/density on segregation indices in order to address this. Or perhaps introduce some spatial indices as I suggested in the previous comment.


```{r}
rm(list = ls()); gc()

library(tidyverse)

load("DB nedu zona censal mayores25.RData")
load("matriz indices seg mayores25.RData")
```

```{r}
res <- as.data.frame(res)
res$codigo_comuna <- levels(as.factor(educ$codigo_comuna))

pop <- educ %>% 
  dplyr::select(codigo_comuna, total, area_km2) %>% 
  group_by(codigo_comuna) %>% 
  summarise(total = sum(total),
            area_km2 = sum(area_km2))

res <- res %>% 
  full_join(pop) %>% 
  mutate(densidad = total/area_km2)

rm(pop, educ)

colnames(res)[1:20] <- c(  # Uneveness
     "Dissimilarity",
     "Giny", 
     "Entropy", 
     "Atkinson (b = 0.1)",
     "Atkinson (b = 0.5)",
     "Atkinson (b = 0.9)",
  # Exposure
     "Interaction",
     "Isolation",
     "Correlation", 
  # Concentration
     "Delta", 
     "Absolute concentration",
     "Relative concentration",
  # Centralization
     "Proportion in central city", 
     "Absolute centralization", 
     "Relative centralization", 
  # Clustering
     "Absolute clustering", 
     "Spatial proximity", 
     "Relative clustering", 
     "Distance decay interaction",
     "Distance decay isolation" 
)

```

```{r}
my_cors <- lapply(1:20, function(x) cor(res[,x], res[, "total"], use = "pairwise.complete.obs"))

my_lms <- lapply(1:20, function(x) lm(res[,x] ~ res[, "total"]))

summaries <- lapply(my_lms, summary)

mods <- data.frame(indice = colnames(res)[1:20],
                   cor_pob = unlist(my_cors),
                   coef_pob = sapply(my_lms, coef)[2,],
                   pval_pob = sapply(summaries, function(x) x$coefficients[2,4]))

mods$test_pob <- mods$pval < 0.05
```

```{r}
my_cors <- lapply(1:20, function(x) cor(res[,x], res[, "densidad"], use = "pairwise.complete.obs"))

my_lms <- lapply(1:20, function(x) lm(res[,x] ~ res[, "densidad"]))

summaries <- lapply(my_lms, summary)

mods$cor_dens <- unlist(my_cors)
mods$coef_dens <- sapply(my_lms, coef)[2,]
mods$pval_dens <- sapply(summaries, function(x) x$coefficients[2,4])
mods$test_dens <- mods$pval_dens < 0.05

rm(summaries, my_lms, my_cors)
```

```{r fig.height=8.5, fig.width=6}
library(scales)
library(ggpubr)

ggarrange(
  
res %>% 
  pivot_longer(cols = Dissimilarity:`Distance decay isolation`,
               names_to = "indice",
               values_to = "indice_valor") %>% 
  ggplot(aes(x = total, y = indice_valor)) + 
  geom_point() +
  geom_label(data = mods, aes(label = round(cor_pob, 2), y = Inf, x = Inf), 
             vjust = 1, hjust = 1, size = 2.5) +
  scale_x_continuous(labels = comma, n.breaks = 4) +
  theme_minimal(base_size = 8) +
  theme(axis.text.x = element_text(angle = 45, vjust = .9)) +
  facet_wrap("indice", scales = "free_y") +
  labs(title = "Correlation between segregation indices and unit populaton size/density",
       x = "Population size",
       y = "Segregation index")
,

res %>% 
  pivot_longer(cols = Dissimilarity:`Distance decay isolation`,
               names_to = "indice",
               values_to = "indice_valor") %>% 
  ggplot(aes(x = densidad, y = indice_valor)) + 
  geom_point() +
  geom_label(data = mods, aes(label = round(cor_dens, 2), y = Inf, x = Inf), 
             vjust = 1, hjust = 1, size = 2.5) +
  scale_x_continuous(labels = comma, n.breaks = 4) +
  theme_minimal(base_size = 8) +
  theme(axis.text.x = element_text(angle = 45, vjust = .9)) +
  facet_wrap("indice", scales = "free_y") +
  labs(title = "",
       x = "Population density",
       y = "Segregation index")
,

ncol = 1, nrow = 2
)

```

